\bfseries\large Phosphatidylinositol 4,5-bisphosphate regulates cilium transition zone maturation in \textit{Drosophila melanogaster}

### 2018-02-26 latex/paper.tex
Various edits

We previously described the presence of occasional triplet microtubules
in electron micrographs of axonemes from \sigd{} \citep{wei2008depletion}.
While centrioles and basal bodies contain microtubule triplets,
restriction of the C-tubule at the TZ causes the axoneme to
consist of microtubule doublets rather than triplets throughout its length.
Consistent with a defect in TZ barrier formation and presence of
microtubule triplets in axonemes,
we observed that a subset of cilia in \sigd{}
contained puncta of Ana1, a core centriolar protein, at the tips
of TZs marked with Cep290.
This TZ-distal Ana1 localization is observed at a relatively low penetrance
and high intra-cyst correlation.
We found that treatment of germ cells with Taxol as in \citep{riparbelli2013unique}
could increase the penetrance of this phenotype.
Taxol treatment stabilizes microtubules and is thought to disrupt TZ maturation
by interrupting microtubule rearrangements, and similarly induces much longer cilia
\citep{riparbelli2012assembly}.
Taxol treatment alone did not induce this phenotype in wild-type controls in our hands
at varying concentrations ($p$ <0.01 with 5\% penetrance),
suggesting that loss of \PIP{} acts upstream of or prior to microtubule rearrangement.
Triplet microtubules within the axoneme are also observed in Taxol-treated cells \citep{riparbelli2013unique}.
It is possible that the TZ defect allows C-tubules and some centriolar proteins
to migrate to the distal end.
We did not observe this behaviour with Asterless (CEP152)
\citep{dzhindzhev2010asterless, blachon2008drosophila},
another centriolar protein ($p$ <0.01), suggesting that these TZ-distal
sites are not fully centriolar in protein composition.

@@

We found that the type I phosphatidylinositol 4-phosphate kinase (PIPKI) Skittles
suppresses TZ elongation in \sigd{} cells in a cilium-autonomous
manner.
We also found that cells mutant for \textit{sktl} contained hyperelongated
TZs similar to SigD-expressing cells.
Sktl also localizes to centrioles,
and our results suggest that Skittles acts locally at cilia to
regulate TZ length.
PIPKI$\gamma$ is linked to lethal congenital contractural syndrome
type 3 (LCCS3).
It has previously been suggested that LCCS3 might represent a ciliopathy.
PIPKI$\gamma$ is linked to LCCS3 with features reminiscent
of ciliopathies such as micrognathia, and a role for the LCCS1 protein
Gle3 in cilia.
In addition, all three human PIPKI isoforms are highly expressed in testes
and PIPKI$\alpha$ and PIPKI$\beta$ are important for fertility in mice
\citep{hasegawa2012phosphatidylinositol}.

@@

%%
%    FIGURE 1
%
\textbf{Figure 1. SigD expression induces transition zone hyperelongation.}
\begin{enumerate}[label={(\Alph*)}, nolistsep]
\item \sigd{} expression induces Cep290 hyperelongation in mature cilia.
  Stages correspond to descriptions in \citep{cenci1994chromatin}.
\item Paired Ana1-Cep290 lengths in nascent and mature cilia ($n$ \textgreater 30).
\item \sigd{} expression induces hyperelongation of other TZ markers, such as Cby, Mks
  and acetylated tubulin.
\item TZ hyperelongation persists through meiosis but does not affect
  axoneme outgrowth.
\end{enumerate}

>>>>

We previously described the presence of occasional triplet microtubules
in electron micrographs of axonemes from \sigd{} \citep{wei2008depletion}.
While centrioles and basal bodies contain microtubule triplets,
restriction of the C-tubule at the TZ causes the axoneme to
consist of microtubule doublets rather than triplets throughout its length.
Consistent with a defect in TZ barrier formation and presence of
microtubule triplets in axonemes,
we observed that a subset of cilia in \sigd{}
contained puncta of Ana1, a core centriolar protein, at the tips
of TZs marked with Cep290.
This TZ-distal Ana1 localization is observed at a relatively low penetrance
and high intra-cyst correlation.
We found that treatment of germ cells with Taxol as in \citep{riparbelli2013unique}
could increase the penetrance of this phenotype.
Taxol treatment stabilizes microtubules and is thought to disrupt TZ maturation
by interrupting microtubule rearrangements, and similarly induces much longer cilia
\citep{riparbelli2012assembly}.
Taxol treatment alone did not induce this phenotype in wild-type controls in our hands
at varying concentrations ($p$ <0.01 with 5\% penetrance),
suggesting that loss of \PIP{} acts upstream of or prior to microtubule rearrangement.
Triplet microtubules within the axoneme are also observed in Taxol-treated cells \citep{riparbelli2013unique}.
It is possible that the TZ defect allows C-tubules and some centriolar proteins
to migrate to the distal end.
We did not observe this behaviour with Asterless (CEP152)
\citep{dzhindzhev2010asterless, blachon2008drosophila} ($p$ <0.01),
another centriolar protein, suggesting that these TZ-distal
sites are not fully centriolar in protein composition.

@@

We found that the type I phosphatidylinositol 4-phosphate kinase (PIPKI) Skittles
suppresses TZ elongation in \sigd{} cells in a cilium-autonomous
manner.
We also found that cells mutant for \textit{sktl} contained hyperelongated
TZs similar to SigD-expressing cells.
Sktl also localizes to centrioles,
and our results suggest that Skittles acts locally at cilia to
regulate TZ length.
PIPKI$\gamma$ is linked to lethal congenital contractural syndrome
type 3 (LCCS3).
It has previously been suggested that LCCS3 might represent a ciliopathy.
PIPKI$\gamma$ is linked to LCCS3 with features reminiscent
of ciliopathies such as micrognathia, and a role for the LCCS1 protein
Gle1 in cilia \citep{}.
In addition, all three human PIPKI isoforms are highly expressed in testes
and PIPKI$\alpha$ and PIPKI$\beta$ are important for fertility in mice
\citep{hasegawa2012phosphatidylinositol}.

@@

%%
%    FIGURE 1
%
\textbf{Figure 1. SigD expression induces transition zone hyperelongation.}
\begin{enumerate}[label={(\Alph*)}, nolistsep]
\item \sigd{} expression induces Cep290 hyperelongation in mature cilia.
  Stages correspond to descriptions in \citep{cenci1994chromatin}.
\item Paired Ana1-Cep290 lengths in nascent and mature cilia.
\item \sigd{} expression induces hyperelongation of other TZ markers, such as Cby, Mks
  and acetylated tubulin.
\item TZ hyperelongation persists through meiosis but does not affect
  axoneme outgrowth.
\end{enumerate}
